# Peer Assessment 1
Gyubok Lee

```{r setoptions, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data

Our dataset is [Activity monitoring data]("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip") and the variables in this data are such:

* steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)

* date: The date on which the measurement was taken in YYYY-MM-DD format

* interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

First, I downloaded the file and I will load my data.


## Download, loading and preprocessing the data

Note: I have my working directory on my desktop
```{r}
setwd("~/Desktop")
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile="~/Desktop/activity.zip",method="curl")
unzip("~/Desktop/activity.zip")
activity <- read.csv("~/Desktop/activity.csv")
```


## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day

```{r, result="asis"}
steps <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

Histogram of the total number of steps taken each day

```{r}
hist(steps, main="Histogram of the total number of steps taken per day")
```

Mean of the total number of steps taken per day

```{r}
mean(steps) 
```

Median of the total number of steps taken per day

```{r}
median(steps) 
```

## What is the average daily activity pattern?

Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days.

```{r, result="asis"}
aversteps <- tapply(activity$steps, activity$interval, mean, na.rm=TRUE)
interval <- unique(activity$interval)
plot(interval, aversteps, type="l", xlab= "5-Minute Interval", ylab="Aver. Number of Steps Taken")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
interval <- sprintf("%04d", interval)
interval <- format(strptime(interval, format="%H%M"), format = "%H:%M")
interval[which(aversteps==max(aversteps))]
```


## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
missing <- dim(activity) - dim(activity[complete.cases(activity), ])
missing[1]
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

=> It would not be the best way, but I consider using the mean of data as missing values is a quite decent approximation of missing values. Since using the median might treat all missing values zeros, I found using the mean is the most convenient yet efficient way for now.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity1 <- activity
activity1[is.na(activity1)] <- (mean(activity$steps, na.rm=TRUE))
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
steps1 <- tapply(activity1$steps, activity$date, sum)
hist(steps1, main="Histogram of the total number of steps taken per day")

mean(activity$steps, na.rm=TRUE)
median(activity$steps, na.rm=TRUE)

mean(activity1$steps, na.rm=TRUE)
median(activity1$steps, na.rm=TRUE)
```

Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

=> There seems no difference by just looking at the mean and the median. The impact is that we do not need to worry about missing values when calculating statstics that we want. Since missing values are replaced by the mean, the first and the second part will not vary much.

# Are there differences in activity patterns between weekdays and weekends?

Sepereate weekend and weekday data
```{r}
activity1$date <- weekdays(as.Date(activity$date, "%Y-%m-%d"))

weekend = data.frame()
weekday = data.frame()
for (i in 1:dim(activity1)[1]) {
        if (activity1$date[i]=="Saterday"||activity1$date[i]=="Sunday"){
                weekend <- rbind(weekend, activity1[i,])}
        else {weekday <- rbind(weekday, activity1[i,])}}
```

Declare factors 
```{r}
weekendmean <- tapply(weekend$steps, weekend$interval, mean)
weekdaymean <- tapply(weekday$steps, weekday$interval, mean)
interval <- unique(activity1$interval)
```

Gather all factors in one data to graph
```{r}
data <- data.frame(c(weekendmean,weekdaymean))
finaldata <- cbind(rep(interval, 2), data, rep(c("weekend","weekday"), each = 288))
colnames(finaldata) <- c("interval", "steps", "day")
```

Graph the data
```{r}
library(lattice)
xyplot(finaldata$steps ~ finaldata$interval |finaldata$day, data = mtcars, pch=19, xlab="Interval",  ylab="Number of Steps",layout=c(1,2),type="l")
```




